### 2013-04-09 paper/paper.tex
Re-add sections that were commented-out

\documentclass[conference]{IEEEtran}
\usepackage{fancyhdr}
\usepackage[pdftex]{graphicx}
\graphicspath{{./figures/}}
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\usepackage[cmex10]{amsmath}
\usepackage{algorithm,algpseudocode}
\usepackage{hyperref} % must come before url
\usepackage{url}
\usepackage{color} % for \textcolor
\usepackage{subfig} % for \subfloat

\newcommand{\TODO}[1]{
  {\Large \textcolor{red}{\textbf{TODO: }#1}}
}

\begin{document}

% can use linebreaks \\ within to get better formatting as desired
\title{Cost- and Deadline-Constrained Provisioning for Scientific Workflow
Ensembles in IaaS Clouds}

% IEEE Format
\author{\IEEEauthorblockN{Maciej
Malawski\IEEEauthorrefmark{1}\IEEEauthorrefmark{3}, Gideon
Juve\IEEEauthorrefmark{2}, Ewa Deelman\IEEEauthorrefmark{2} and Jarek
Nabrzyski\IEEEauthorrefmark{3}} 
\IEEEauthorblockA{\IEEEauthorrefmark{1}AGH University of Science and
Technology, Dept. of Computer Science, Krakow, Poland, Email:
malawski@agh.edu.pl} 
\IEEEauthorblockA{\IEEEauthorrefmark{2}USC Information Sciences Institute, 4676
Admiralty Way, Marina del Rey, CA, USA, Email: \{gideon,deelman\}@isi.edu} 
\IEEEauthorblockA{\IEEEauthorrefmark{3}University of Notre Dame, Center for
Research Computing, 111 ITC, Notre Dame, IN, USA, Email: naber@nd.edu}
}

% ACM Format
%\numberofauthors{3}
%\author{
%    \alignauthor Maciej Malawski\\
%       \affaddr{AGH University of Science and Technology}\\
%       \affaddr{Dept. of Computer Science}\\
%       \affaddr{Krakow, Poland; and}\\
%       \affaddr{University of Notre Dame}\\
%       \affaddr{Center for Research Computing}\\
%       \affaddr{111 ITC, Notre Dame, IN, USA}\\
%       \email{mmalawsk@nd.edu}
%    \alignauthor Gideon Juve and Ewa Deelman\\
%       \affaddr{USC Information Sciences Institute}\\
%       \affaddr{4676 Admiralty Way}\\
%       \affaddr{Marina del Rey, CA, USA}\\
%       \email{\{gideon,deelman\}@isi.edu}
%    \alignauthor Jarek Nabrzyski\\
%       \affaddr{University of Notre Dame}\\
%       \affaddr{Center for Research Computing}\\
%       \affaddr{111 ITC, Notre Dame, IN, USA}\\
%       \email{naber@nd.edu}
%}

\maketitle

\thispagestyle{fancy}
\lhead{}
\rhead{}
\chead{}
\lfoot{\footnotesize{
SC12, November 10-16, 2012, Salt Lake City, Utah, USA
\newline
978-1-4673-0806-9/12/\$31.00 \copyright 2012 IEEE}}
\rfoot{}
\cfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{abstract}

Large-scale applications expressed as scientific workflows are
often grouped into ensembles of inter-related workflows. In this paper, we
address a new and important problem concerning the efficient management of
such ensembles under budget and deadline constraints on Infrastructure-
as-a-Service (IaaS) clouds. We discuss, develop, and assess algorithms based
on static and dynamic strategies for both task scheduling and resource
provisioning. We perform the evaluation via simulation using a set of
scientific workflow ensembles with a broad range of budget and deadline
parameters, taking into account uncertainties in task runtime estimations,
provisioning delays, and failures. We find that the key factor determining the
performance of an algorithm is its ability to decide which workflows in an
ensemble to admit or reject for execution. Our results show that an admission
procedure based on workflow structure and estimates of task runtimes can
significantly improve the quality of solutions.

\end{abstract}

% IEEE does not have these
%\category{H.3.4}{Systems and Software}{Distributed systems}
%\category{D.4.8}{Performance}{Simulation}
%\terms{DAG, Scheduling, Simulation}
%\keywords{Scientific workflows, Workflow ensembles, Cloud computing, Infrastructure-as-a-Service}

\section{Introduction}
\label{sec:intro}

Scientific workflows, usually represented as Directed Acyclic Graphs (DAGs),
are an important class of applications that lead to challenging problems in
resource management on grid and utility computing systems. Workflows for large
computational problems are often composed of several inter-related workflows
grouped into {\em ensembles}.  Workflows in an ensemble typically have a
similar structure, but they differ in their input data, number of tasks, and
individual task sizes.

There are many applications that require scientific workflow ensembles.
CyberShake~\cite{Callaghan2011}, for example, uses ensembles to generate
seismic hazard maps.  Each workflow in a CyberShake ensemble generates a
hazard curve for a particular geographic location, and several hazard curves
are combined to create a hazard map.  In 2009 CyberShake was used to generate
a map that required an ensemble of 239 workflows.  Similarly, users of
Montage~\cite{Deelman2008} often need several workflows  with different
parameters to generate a set of image mosaics that can be combined  into a
single, large mosaic. The Galactic Plane ensemble, which generates several
mosaics of the  entire sky in different wavelengths, consists of 17 workflows,
each of which contains 900 sub-workflows.  Another ensemble example is the
Periodograms application~\cite{Vockler2011}, which searches for extrasolar
planets by detecting periodic dips in the light intensity of their host star.
Due to the large scale of the input data, this application is often split up
into multiple batches processed by different workflows. Additional workflows
are created to run the analysis using different parameters. A recent analysis
of Kepler satellite data required three ensembles of 15 workflows.

Workflows in an ensemble may differ not only in their parameters, but also in
their priority. For example, in CyberShake some geographic locations may be in heavily
populated areas or in strategic locations such as power plants, while others
may be less important. Scientists typically prioritize the workflows in such
an ensemble so that important workflows are finished first. This enables them
to see critical results early, and helps them to choose the most important
workflows when the time and financial resources available for computing are
limited.

Infrastructure-as-a-Service (IaaS) clouds offer the ability to provision 
resources on-demand according to a pay-per-use model. These systems 
are regarded by the scientific community as a potentially attractive source of 
low-cost computing resources~\cite{Ostermann2010, Keahey2009}. In contrast 
to clusters and grids, which typically offer best-effort quality of service, clouds 
give more flexibility in creating a controlled and managed computing environment.
Clouds provide the ability to adjust resource capacity according to the changing
demands of the application, often called auto-scaling. However, giving users 
more control also requires the development of new methods for task 
scheduling and resource provisioning. Resource management decisions required 
in cloud scenarios not only have to take into account performance-related 
metrics such as workflow makespan or resource utilization, but must also consider
budget constraints, since the resources from commercial clouds
usually have monetary costs associated with them~\cite{Durkee2010}.

In this paper, we aim to gain insight into resource management challenges when
executing scientific workflow ensembles on clouds. We address a
new and important problem of maximizing the number of completed workflows from
an ensemble under both budget and deadline constraints. The motivation for this
work is to answer the fundamental question of concern to a researcher: How much
computation can be completed given the limited budget and timeframe of a
research project?

The main contributions of this paper are:
\begin{itemize}
  \item we define the problem of scheduling prioritized workflow ensembles
  under budget and deadline constraints,
  \item we analyze and develop several dynamic (online) and static (offline) 
  algorithms for task scheduling and resource provisioning that rely on workflow 
  structure information (critical paths and workflow levels) and estimates of 
  task runtimes,
  \item we evaluate these algorithms using a simulator based on 
  CloudSim~\cite{Calheiros2011}, which models the infrastructure and the 
  application,taking into account uncertainties in task runtime estimates, 
  provisioning delays, and failures, 
  \item we discuss the performance of the algorithms on a set of
  synthetic workflow ensembles based on important, real scientific
  applications, using a broad range of different application scenarios and
  varying constraint values.
\end{itemize}

The paper is organized as follows: after an analysis of related work in
Section~\ref{sec:related}, Section~\ref{sec:problem} describes the
infrastructure and application model we are targeting in this paper. In
Section~\ref{sec:algorithms} we describe the dynamic and static algorithms we
developed. Section~\ref{sec:performance} presents test scenarios and performance
metrics, while Section~\ref{sec:results} discusses the results of simulation
studies of these algorithms. Finally, general conclusions, lessons learned, and
future work are outlined in Section~\ref{sec:conclusions}.

\section{Related Work}
\label{sec:related}
General policy and rule-based approaches to dynamic provisioning (e.g. Amazon
Auto Scaling~\cite{Autoscaling} and
RightScale~\cite{RightScale}) allow the size
of a resource pool to be adjusted based on infrastructure and
application metrics. A typical infrastruc\-ture-specific metric is system load,
whereas application-speci\-fic metrics include response time and length of a task or
of a request queue. It is possible to set thresholds and limits to tune the behavior
of these autoscaling systems, but no support for complex applications is provided.
Policy-based approaches for scientific workloads (e.g. \cite{Marshall2010,
Kim2011}) also allow to scale the cloud resource pool or to extend the
capabilities of clusters using cloud-burst techniques. Our approach is different
in that we consider workflows, while policy based approaches typically consider
bags of independent tasks or unpredictable batch workloads. This enables us to
take advantage of workflow-aware heuristics that cannot be applied to
independent tasks.

Our work is related to the strategies for deadline-con\-strained
cost-minimization workflow scheduling, developed for utility grid systems. However, our problem is
different from \cite{Yu2005} and \cite{Abrishami2010} in that we consider ensembles of workflows in
IaaS clouds, which allow one to provision resources on a per-hour billing model,
rather than utility grids, which allow one to choose from a pool of existing
resources with a per-job billing model. Our work is also different from
cloud-targeted autoscaling solution~\cite{Mao2011} in that we consider ensembles
of workflows rather than unpredictable workloads containing workflows. We also consider budget constraints
rather than cost minimization as a goal. In other words, we assume that there is
more work to be done than the available budget, so some work must be rejected.
Therefore, cost is not something we optimize (i.e. an objective), but rather a constraint.

This work is related to bi-criteria scheduling and multi-criteria scheduling of
workflows~\cite{Wieczorek2009,Prodan2010,Dongarra2007}. These approaches are
similar to ours in that we have two scheduling criteria: cost and makespan. The
challenge in multi-criteria scheduling is to derive an objective function that
takes into account all of the criteria. In our case one objective (amount
of work completed) is subject to optimization, whereas time and cost are
treated as constraints. Other approaches~\cite{Talukder2009,Pandey2010} use
metaheuristics that usually run for a long time before producing good results,
which makes them less useful in the scenarios we consider in this paper. Our work can also
be regarded as an extension of the budget-constrained workflow scheduling
\cite{Sakellariou2007} in the sense that we are dealing with workflow ensembles
and the deadline constraint is added.

\section{Problem Description}
\label{sec:problem}

\underline{Resource Model}  We assume a resource model similar to Amazon's Elastic Compute Cloud (EC2),
where virtual machine (VM) instances may be provisioned on-demand and are billed by
the hour, with partial hours being rounded up. Although there may be
heterogeneous VM types with different amounts of CPU, memory, disk space, and
I/O, for this paper we focus on a single VM type because we assume that for most
applications there will typically be only one or two VM types with the best
price/performance ratio for the application~\cite{Juve2009}. We assume that
a submitted task has exclusive access to a VM instance and that there is no
preemption. We also assume that there is a delay between the time that a new VM
instance is requested and when it becomes available to execute tasks.

\underline{Application Model} The target applications are ensembles of
scientific workflows that can be modeled as DAGs, where the nodes in the graph
represent computational tasks, and the edges represent data- or control-flow
dependencies between the tasks. We assume that runtime estimates for the
workflow tasks are known, but that they are not perfect and may vary based on
a uniform distribution of $\pm p\%$.

This study uses synthetic workflows that were generated using historical data
from real applications~\cite{Bharathi2008}. The applications come from a wide
variety of domains including: bioinformatics (Epigenomics, SIPHT: sRNA
identification protocol using high-throughput technology), astronomy
(Montage), earthquake science (CyberShake), and physics (LIGO). The synthetic
workflows were generated using code developed in~\cite{WorkflowGenerator},
with task runtimes based on distributions gathered from running real
workflows.

Although workflows are often data-intensive, the algorithms described
here do not currently consider the size of input and output data
when scheduling tasks. Instead, it is assumed that all workflow data is stored in
a shared cloud storage system, such as Amazon S3, and that intermediate data
transfer times are included in task runtimes. It is also assumed that
data transfer times between the shared storage and the VMs are equal for
different VMs so that task placement decisions do not impact the
runtime of the tasks.

We assume that each workflow in an ensemble is given a numeric priority that
indicates how important the workflow is to the user. As such, the priorities
indicate the utility function of the user. These priorities are absolute in
the sense that completing a workflow with a given priority is more valuable
than completing all other workflows in the ensemble with lower priorities
combined. The goal of the workflow ensemble scheduling and cloud provisioning
problem is to complete as many high-priority workflows as possible given a
fixed budget and deadline. Only workflows for which all tasks are finished by
the deadline are considered to be complete---partial results are not usable in
this model.

\underline{Performance Metric}
\label{sec:perf_metric}
In order to precisely define the objective of the algrithms it is necessary to
introduce a metric that can be used to score the performance of the different
algorithms on a given problem (ensemble, budget, and deadline). The simplest
approach is to count the number of workflows in the ensemble that each algorithm
is able to complete within the budget before the deadline, but this metric does
not account for the priority-based utility function specified by the user. Using
the counting approach, a less efficient algorithm may be able to complete a
large number of low-priority workflows by executing the smallest workflows
first. In order to account for the priority, we use an exponential
scoring defined as:
$$ Score(e) = \sum_{w~\in~Completed(e)}{2^{-Priority(w)}} $$
where $Completed(e)$ is the set of workflows in ensemble $e$ that was completed
by the algorithm, and $Priority(w)$ is the priority of workflow $w$ such that
the highest-priority workflow has $Priority(w)=0$, the next highest workflow has
$Priority(w)=1$, and so on. This exponential scoring function gives the highest
priority workflow a score that is higher than all the lower- priority workflows
combined:
$$ 2^{-p} > \sum_{i~=~p+1,~\ldots}2^{-i} $$
This scoring is consistent with our definition of the problem, which is to
complete as many workflows as possible, according to their priorities, given a
set budget and deadline.

\section{Algorithms}
\label{sec:algorithms}

This section describes three algorithms that were developed to schedule and
provision resources for ensembles of workflows on the cloud under budget and
deadline constraints.

\begin{algorithm}[tb]
\caption{Dynamic provisioning algorithm for DPDS}
\label{alg:prov}
{\footnotesize
\begin{algorithmic}[1]
\Require $c$: consumed budget; $b$: total budget; $d$: deadline; $p$: price;
$t$: current time; $u_h$: upper utilization threshold; $u_l$: lower utilization
threshold; $v_{max}$: maximum number of VMs
\Procedure{Provision}{}
  \State $V_R\gets$ set of running VMs
    \State $V_C\gets$ set of VMs completing billing cycle
    \State $V_T\gets \emptyset$ \Comment{set of VMs to terminate}
    \State $n_T\gets 0$ \Comment{number of VMs to terminate}
    \If{$ b - c < |V_C| * p\ or\ t > d $ }
      \State $n_T\gets |V_R| - \lfloor(b-c)/p\rfloor$
      \State $V_T\gets$ select $n_T$ VMs to terminate from $V_C$
      \State \Call{Terminate}{$V_T$} \label{l:terminate1}
    \Else 
    \State $u\gets$ current VM utilization
      \If{$u>u_h$ and $|V_R| < v_{max}*N_{VM}$}
        \State \Call{Start}{$new\ VM$}
      \ElsIf{$u<u_l$}
        \State $V_I\gets$ set of idle VMs
        \State $n_T\gets \lceil|V_I|/2\rceil$ \label{l:nT2}
      \State $V_T\gets$ select $n_T$ VMs to terminate from $V_I$
        \State \Call{Terminate}{$V_T$} \label{l:terminate2}
      \EndIf 
    \EndIf
\EndProcedure
\end{algorithmic}}
\end{algorithm}

\subsection{Dynamic Provisioning Dynamic Scheduling (DPDS)}
\label{sec:dpds}

DPDS is an online algorithm that provisions resources and schedules tasks at
runtime. It consists of two main parts: a provisioning procedure, and a
scheduling procedure.

DPDS' provisioning procedure is based on resource utilization. DPDS starts with
a fixed number of resources calculated based on the available time and budget,
and adjusts the number of resources according to how well they are utilized by
the application.
Given a budget in dollars $b$, deadline in hours $d$, and the hourly price of a
VM in dollars $p$, it is possible to calculate the number of VMs, $N_{VM}$, to
provision so that the entire budget is consumed before the deadline:
%
\begin{equation}
\label{eq:static-plan}
N_{VM} = \lceil b / (d * p) \rceil
\end{equation}
%
DPDS provisions $N_{VM}$ VMs at the start of the ensemble execution, then it
periodically computes resource utilization using the percentage of idle VMs
over time and adjusts the number of VMs if the utilization is above or below
given thresholds. Because it is assumed that VMs are billed by the hour, DPDS
only considers VMs that are approaching their hourly billing cycle when
deciding which VMs to terminate. This dynamic provisioning algorithm is shown
in Algorithm~\ref{alg:prov}.

\begin{algorithm}[tb]
\caption{Priority-based scheduling algorithm for DPDS}
\label{alg:ds}
{\footnotesize
\begin{algorithmic}[1]
\Procedure{Schedule}{}
    \State $P\gets$ empty priority queue
  \State $IdleVMs\gets$ set of idle VMs
  \For{root task $t$ in all workflows}
      \State \Call{Insert}{$t,P$} 
    \EndFor
    \While{deadline not reached}
      \While{$IdleVMs \neq \emptyset\ and\ P \neq \emptyset$}
        \State $v\gets$ \Call{SelectRandom}{$IdleVMs$}
        \State $t\gets$ \Call{Pop}{$P$}
        \State \Call{Submit}{$t,v$}
      \EndWhile
      \State Wait for task $t$ to finish on VM $v$
      \State Update $P$ with ready children of $t$
    \State \Call{Insert}{$v,IdleVMs$}
    \EndWhile
\EndProcedure
\end{algorithmic}
}
\end{algorithm}

The set of VMs completing their billing cycle is determined by both
the provisioner interval, and the termination delay of the provider. This
guarantees that VMs can be terminated before they start the next billing cycle
and prevents the budget from being overrun. The VMs terminated in line
\ref{l:terminate1} of Algorithm~\ref{alg:prov} are the ones that would overrun
the budget if not terminated in the current provisioning cycle. The VMs
terminated in line \ref{l:terminate2} are chosen to increase the resource
utilization to the desired threshold. In order to prevent instances that have
already been paid for from being terminated too quickly, no more than half of
the idle resources are terminated during each provisioning cycle. To avoid an
uncontrolled increase in the number of instances, which may happen in the case
of highly parallel workflows, the provisioner will not start a new VM if the
number of running VMs is greater than the product of $N_{VM}$ (from
Equation~\ref{eq:static-plan}) and an autoscaling parameter, $v_{max}$. Unless
otherwise specified, $v_{max}$ is assumed to be 1.

In order to schedule individual workflow tasks onto available VMs, DPDS uses the
dynamic, priority-based scheduling procedure shown in Algorithm~\ref{alg:ds}.
Initially, the ready tasks from all workflows in the ensemble are added to a
priority queue based on the priority of the workflow to which they belong. If
there are idle VMs available, and the priority queue is not empty, the next task
from the priority queue is submitted to an arbitrarily chosen idle VM. The
process is repeated until there are no idle VMs or the priority queue is empty.
The scheduler then waits for a task to finish, adds its ready children to the
priority queue, marks the VM as idle, and the entire process repeats until the
deadline is reached.

DPDS guarantees that tasks from lower priority workflows are always deferred
when higher-priority tasks are available, but lower-priority tasks can still
occupy idle VMs when higher-priority tasks are not available. Since there is
no preemption, long-running low-priority tasks may delay the execution of
higher-priority tasks. In addition, tasks from low priority workflows may be
executed even though there is no chance that those workflows will be completed
within the current budget and deadline. Fig.~\ref{fig:algorithm-example-dpds}
shows an example schedule generated using the DPDS algorithm. The figure
illustrates how tasks from lower priority workflows backfill idle VMs when
tasks from higher priority workflows are not available.

\subsection{Workflow-Aware DPDS (WA-DPDS)}

DPDS does not use any information about the structure of the
workflows in the ensemble when scheduling tasks.  It does not
consider whether a lower priority task belongs to a workflow that will never be
able to complete given the current budget and deadline. As a result, DPDS may
start lower priority tasks just to keep VMs busy that will end up delaying
higher priority tasks later on, making it less likely that higher priority
workflows will be able to finish.

The Workflow-Aware DPDS (WA-DPDS) algorithm extends DPDS by introducing a
workflow admission procedure, which is invoked whenever WA-DPDS sees the first
task of a new workflow at the head of the priority queue (i.e. when no other
tasks from the workflow have been scheduled yet). The admission procedure---
shown in Algorithm~\ref{alg:wa-dpds}---estimates whether there is enough
budget remaining to admit the new workflow; if there is not, then the workflow
is rejected and its tasks are removed from the queue. WA-DPDS compares the
current cost (consumed budget) and remaining budget, taking into account the
cost of currently running VMs, and the cost of workflows that have already
been admitted. In addition, it adds a small safety margin of \$0.10 (10\% of the
compute hour cost) to avoid going over budget. We found the admission procedure useful not only to prevent
low-priority workflows from delaying high-priority ones, but also to reject
large and costly workflows that would overrun the budget and admit smaller
workflows that can efficiently utilize idle resources in 
ensembles containing workflows of non-uniform sizes. It would also be possible 
to extend this admission procedure to check other constraints, such as whether
the estimated critical path of the new workflow exceeds the time remaining
until the deadline.

\begin{algorithm}[tb]
\caption{Workflow admission algorithm for WA-DPDS}
\label{alg:wa-dpds}
{\footnotesize
\begin{algorithmic}[1]
\Require $w$: workflow; $b$: budget; $c$: current cost
\Procedure{Admit}{$w,b,c$}
    \State $r_n\gets b-c$ \Comment{Budget remaining for new VMs}
    \State $r_c\gets $cost committed to VMs that are running
    \State $r_a\gets $cost to complete workflows previously admitted
  \State $r_m\gets 0.1$ \Comment{Safety margin}
  \State $r_b\gets r_n+r_c-r_a-r_m$ \Comment{Budget remaining}
  \State $c_w\gets$ \Call{EstimateCost}{w}
    \If{$c_w<r_b$}  \textbf{return} $TRUE$
    \Else ~  \textbf{return} $FALSE$
  \EndIf       
\EndProcedure
\end{algorithmic}
}
\end{algorithm}

\subsection{Static Provisioning Static Scheduling (SPSS)}

The previous dynamic (online) algorithms make provisioning and scheduling
decisions at runtime. By contrast, the SPSS algorithm creates a provisioning
and scheduling plan before running any workflow tasks. This enables SPSS to
start only those workflows that it knows can be completed given the deadline and
budget constraints, and eliminates any waste that may be allowed by the dynamic
algorithms. However, the disadvantage of SPSS is that it is sensitive to dynamic
changes in the environment (see Sections~\ref{sec:delays} and
\ref{sec:variances}).

The approach used by SPSS is to plan each workflow in the ensemble in priority
order, rejecting any workflows that would exceed the dedline or budget. Once the
plan is complete, the VMs are provisioned and tasks are executed according to
the schedules given by the plan.

The disadvantage of the static planning approach used by SPSS is that it is
sensitive to dynamic changes in the environment and the application that may
disrupt its carefully constructed plan. For example, if there are provisioning
delays, or if the runtime estimates for the tasks are inaccurate, then workflow
execution may diverge from the plan. This issue will be discussed further in
Sections~\ref{sec:delays} and \ref{sec:variances}.

Algorithm~\ref{alg:admit} shows how ensembles are planned in SPSS. Workflows
from the ensemble are considered in priority order. For each workflow, SPSS
attempts to build on top of the current plan by provisioning VMs to schedule the
tasks of the workflow so that it finishes before the deadline with the least
possible cost. If the cost of the new plan is less than the budget, then the new
plan is accepted and the workflow is admitted. If not, then the new plan is
rejected and the process continues with the next workflow in the ensemble. The
idea is that, if each workflow can be completed by the
deadline with the lowest possible cost, then the number of workflows that can be
completed within the given budget will be maximized.

To plan a workflow, the SPSS algorithm assigns sub-deadlines to each individual
task in the workflow, and then schedules each task so as to minimize the cost of
the task while still meeting its assigned sub-deadline. If each task can be
completed by its deadline in the least expensive way, then the cost of the
entire workflow can be minimized without exceeding the deadline. SPSS assigns
sub-deadlines to each task based on the slack time of the workflow, which is
defined as the amount of extra time that a workflow can extend its critical path
and still be completed by the ensemble deadline. For a workflow $w$, the slack
time of $w$ is: $ ST(w) = d - CP(w) $
where $d$ is the deadline and $CP(w)$ is the critical path of $w$. We
assume that $CP(w) \leq d$, otherwise the workflow cannot be completed by the
deadline and must be rejected. 
For large ensembles we expect the critical path
of any individual workflow to be much less than the deadline.

\begin{algorithm}[tb]
\caption{Ensemble planning algorithm for SPSS}
\label{alg:admit}
{\footnotesize
\begin{algorithmic}[1]
\Require $W$: workflow ensemble; $b$: budget; $d$: deadline
\Ensure Schedule as much of $W$ as possible given $b$ and $d$
\Procedure{PlanEnsemble}{$W,b,d$}
    \State $P\gets \emptyset$ \Comment{Current plan}
    \State $A\gets \emptyset$ \Comment{Set of admitted DAGs}
    \For{$w\ in\ W$}
        \State $P^\prime \gets$\ \Call{PlanWorkflow}{$w,P,d$}
        \If{$Cost(P^\prime) \le b$}
            \State $P\gets\ P^\prime$ \Comment{Accept new plan}
            \State $A \gets A\ +\ w$ \Comment{Admit w}
        \EndIf
    \EndFor
    \State \textbf{return} $P,A$
\EndProcedure
\end{algorithmic}
}
\end{algorithm}

A task's level is the length of the longest path between the task and an entry
task of the workflow:
%
$$
Level(t) = \begin{cases}
0,&$if~$Pred(t) = \emptyset\cr
\max_{p \in Pred(t)}Level(p)+1,&$otherwise.$
\end{cases}
$$
%
SPSS distributes the slack time of the workflow by level, so that each level
of the workflow gets a portion of the workflow's slack time proportional to
the number of tasks in the level and the total runtime of tasks in the level.
The idea is that levels containing many tasks and large runtimes should be
given a larger portion of the slack time so that tasks in those levels may be
serialized. Otherwise, many resources need to be allocated to run all of the
tasks in parallel, which may be more costly.

The slack time of a level $l$ in workflow $w$ is given by:
%
$$
ST(l) = ST(w) \Biggl[\left({\alpha}\frac{N(l)}{N(w)}\right) + \left({(1 - \alpha)}\frac{R(l)}{R(w)} \right)\Biggr]
$$
%
where $N(w)$ is the number of tasks in the workflow, $N(l)$ is the number of
tasks in level $l$, $R(w)$ is the total runtime of all tasks in the workflow,
$R(l)$ is the total runtime of all tasks in level $l$, and $\alpha$ is a
parameter between 0 and 1 that causes more slack time to be given to levels
with more tasks (large $\alpha$) or more runtime (small $\alpha$).

The deadline of a task $t$ is then:
%
$$
DL(t) = LST(t) + RT(t) + ST(Level(t))
$$
%
where $Level(t)$ is the level of $t$, $RT(t)$ is the runtime of $t$, and 
$LST(t)$ is the latest start time of $t$ determined by:
%
$$
LST(t) = \begin{cases}
0,&$if~$Pred(t) = \emptyset\cr
\max_{p \in Pred(t)} DL(p),&$otherwise.$
\end{cases}
$$

\begin{algorithm}[tb]
\caption{Workflow planning algorithm for SPSS}
\label{alg:planworkflow}
{\footnotesize
\begin{algorithmic}[1]
\Require $w$: workflow; $P$: current plan; $d$: deadline
\Ensure Create plan for $w$ that minimizes cost and meets deadline $d$
\Procedure{PlanWorkflow}{$w,P,d$}
    \State $P^\prime\gets$ copy of $P$
    \State \Call{DeadlineDistribution}{w,d}
    \For{$t\ in\ w\ sorted\ by\ DL(t)$}
        \State $v \gets$ VM that minimizes cost and start time of t
        \If{$FinishTime(t,v) < DL(t)$}
            \State Schedule(t,v)
        \Else
            \State Provision a new VM v
            \State Schedule(t,v)
        \EndIf
    \EndFor
    \State \textbf{return} $P^\prime$
\EndProcedure
\end{algorithmic}
}
\end{algorithm}

Algorithm~\ref{alg:planworkflow} shows how SPSS creates low-cost plans for each
workflow. The \textsc{PlanWorkflow} procedure first calls
\textsc{DeadlineDistribution} to assign sub-deadlines to tasks.
Then, \textsc{PlanWorkflow} schedules
tasks onto VMs, allocating new VMs when necessary. For each task in the
workflow, the least expensive slot is chosen to schedule the task so that it can
be completed by  its deadline. VMs are allocated in blocks of one billing cycle
(one hour) regardless of the size of the task. When computing the cost of
scheduling a task on a given VM, the algorithm considers idle slots in blocks
that were allocated for previous tasks to be free, while slots in new blocks
cost the full price of a billing cycle. 
For example, if a task has a runtime of
10 minutes, and the price of a block is \$1, then the algorithm will either
schedule the task on an existing VM that has an idle slot larger than 10
minutes for a cost of \$0, or it will allocate a new block on an existing VM,
or provision a new VM, for a cost of \$1. 
If the cost of slots on two different
VMs is equal, then the slot with the earliest start time is chosen. To prevent too
many VMs from being provisioned, the algorithm always prefers to extend the
runtime of existing VMs before allocating new VMs. The result of this is that
the algorithm will only allocate a new block if there are no idle slots on
existing blocks large enough or early enough to complete the task by its
deadline, and it will only allocate a new VM if it cannot add a block to an
existing VM to complete the task by its deadline.

\begin{figure}[tb] 
  \centering
  \subfloat[DPDS]{
  \includegraphics[width=0.40\textwidth]{figures/output-test-DPDS}
  \label{fig:algorithm-example-dpds}
  }\\
  \subfloat[SPSS]{
  \includegraphics[width=0.40\textwidth]{figures/output-test-SPSS}
  \label{fig:algorithm-example-spss}
  }
  \caption[Example schedules generated by the algorithms]{Example
  schedules generated by the algorithms. Each row is a different
  VM. Boxes are tasks colored by workflow.}

\end{figure}

An example schedule generated by SPSS (Fig.~\ref{fig:algorithm-example-spss})
shows how SPSS tends to start many workflows in parallel, running each workflow
over a longer period of time on only a few VMs to minimize cost. In comparison,
the dynamic algorithms tend to run one workflow at a time across many VMs in
parallel.

\section{Evaluation Methods}
\label{sec:performance}

\subsection{Simulator}

To evaluate and compare the three proposed algorithms, we developed a cloud
workflow simulator~\cite{CloudWorkflowSimulator} based on CloudSim~\cite{Calheiros2011}. Our
simulation model consists of Cloud, VM and WorkflowEngine entities. The Cloud entity starts and
terminates VM entities using an Amazon EC2-like API. VM entities simulate the
execution of individual tasks, including randomized variations in runtime. The
WorkflowEngine entity manages the scheduling of tasks and the provisioning of 
VMs based on the chosen algorithm. We assume that the VMs have a single core
and execute tasks sequentially. 
Although CloudSim provides a more advanced infrastructure model which includes
time-sharing and space-sharing policies, we do not use these features since we
are interested mainly in the execution of tasks on VMs and high-level workflow
scheduling and provisioning. 
The simulator reads workflow description files in
a modified version of the DAX format used by the Pegasus Workflow Management
System~\cite{Deelman2005}. 
%The modified file format was created for the
%synthetic workflow generator developed by Bharathi, et al~\cite{Bharathi2008}.

\subsection{Workflow Ensembles}
\label{sec:ensembles}

In order to evaluate the algorithms on a standard set of workflows, we created
randomized ensembles using workflows available from the workflow generator
gallery~\cite{WorkflowGenerator}. The
gallery contains synthetic workflows modeled using structures and parameters
that were taken from real applications. Ensembles were created using synthetic
workflows from five real applications: SIPHT, LIGO, Epigenomics, Montage and
CyberShake. For each application, workflows with 50, 100, 200, 300,
400, 500, 600, 700, 800, 900 and 1000 tasks were created. For each workflow
size, 20 different workflow instances were generated using parameters and task
runtime distributions from real workflow traces. The total collection of
synthetic workflows contains 5 applications, 11 different workflow sizes, and 20
workflow instances, for a total of 1100 synthetic workflows.

Using this collection of workflows, we constructed five different ensemble types:
constant, uniform sorted, uniform unsorted, Pareto sorted and Pareto unsorted.
In the unsorted ensembles, workflows of different sizes are mixed together and
the priorities are assigned randomly.
% so that there is no relationship between
%priority and workflow size. 
For many applications, however, large workflows are
more important to users than small workflows because they represent more
significant computations. To model this, the sorted ensembles are sorted by
size, so that the largest workflows have the highest priority.

Constant ensembles contain workflows that all have the same number of tasks. 
%For each constant ensemble, 
The number of tasks is chosen randomly from the set of
possible workflow sizes. Once the size is determined, then N workflows of that
size are chosen randomly for the ensemble from the set of synthetic workflows.
%for the given application.

Uniform ensembles contain workflows with sizes that are uniformly distributed
among the set of possible sizes. Each workflow 
%in a uniform ensemble 
is selected by first randomly choosing the size of the workflow
% according to a uniform distribution, 
and then randomly choosing a workflow of that size from the set of
synthetic workflows.
% for the given application.

Pareto ensembles contain a small number of larger workflows and a large number
of smaller workflows. Their sizes  
%in a Pareto ensemble 
are chosen according to a Pareto distribution. The distribution was modified so
that the number of large workflows (of size $\geq$ 900) is increased by a small amount to
produce a ``heavy-tail''. This causes Pareto ensembles to have a slightly larger
number of large workflows, which reflects behavior commonly observed in many
computational workloads. 
An example of the distribution of workflow sizes that
occurs in a Pareto ensemble is shown in Fig.~\ref{fig:ensemble-pareto}.

\begin{figure}[ht] 
    \centering
    \includegraphics[width=0.6\columnwidth]{figures/ensemble-pareto}
    \caption[Histogram of workflow sizes in Pareto ensembles]{Histogram of
    workflow sizes in Pareto ensembles. Workflow size is measured in number of
    tasks.}
    \label{fig:ensemble-pareto}
\end{figure}

The number of workflows in an ensemble depends on the particular application,
but we assume that ensemble sizes are on the order of between 10 and 100
workflows, typical of the real applications we have examined (see Section~\ref{sec:intro}). 
For example, the number of
geographical sites of interest to the users of the CyberShake application in
the past has been on the order of 100. 
Second, smaller ensembles consisting of
just a few workflows can be aggregated into a single workflow, so there is no need to
treat them as an ensemble. Similarly, when the number of workflows grows and
each workflow has a large number of tasks, either the deadline and budget
constraints are low enough to prevent many of the workflows from running, or the
problem of efficiently allocating them to the resources becomes similar to a
bag-of-tasks problem, which is easier to solve efficiently.

\subsection{Experimental Parameters}
\label{sec:exp-parameters}

In order to observe the interesting characteristics of the proposed algorithms,
for each ensemble, we selected ranges for deadline and budget that cover a broad
parameter space: from tight constraints, where only a small number of workflows can be completed,
to more liberal constraints where all, or almost all, of the workflows can be
completed. We computed constraint ranges based on
the characteristics of each ensemble. The budget constraints are calculated by
identifying the smallest budget required to execute one of the workflows in the
ensemble (MinBudget), and the smallest budget required to execute all workflows
in the ensemble (MaxBudget):
%
$$
MinBudget = \min_{w\ \in\ e}{Cost(w)}
$$
$$
MaxBudget = \sum_{w\ \in\ e}{Cost(w)}
$$
%
This range---$[MinBudget, MaxBudget]$---is then divided into equal intervals to
determine the budgets to use in each experiment. Similarly, the deadline
constraints are calculated by identifying the smallest amount
of time required to execute a single workflow in the ensemble (MinDeadline),
which is the length of the critical path for the workflow with the shortest
critical path, and by identifying the smallest amount of time required to
execute all workflows (MaxDeadline), which is the sum of the critical paths of
all the workflows:
%
$$
MinDeadline = \min_{w\ \in\ e}{CriticalPath(w)}
$$
$$
MaxDeadline = \sum_{w\ \in\ e}{CriticalPath(w)}
$$
%
This range---$[MinDeadline, MaxDeadline]$---is then divided into equal intervals. 
By computing the budget and deadline constraints in this way we ensure that the
experiments for each ensemble cover the most interesting area of the parameter space for the
ensemble.

In all the experiments we assumed that the VMs have a price of \$1 per VM-hour.
This price was chosen to simplify interpretation of results and should not
affect the relative performance of the different algorithms. In this study the
heterogeneity of the infrastructure is not relevant since we assume that it is
always possible to select a VM type that has the best price to performance ratio
for a given application~\cite{Juve2009}.

All the experiments were run with maximum autoscaling factor ($v_{max}$) set to
1.0 for DPDS and WA-DPDS. After experimenting with DPDS and WA-DPDS we found
that, due to the high parallelism of workflows used, the resource utilization
remains high enough without adjusting the autoscaling rate. Based on experiments
with the target applications, we set the SPSS $\alpha$ parameter for deadline
distribution to be 0.7, which allocates slightly more time to levels with many
tasks.

\section{Discussion of Results}
\label{sec:results}

\subsection{Relative Performance of Algorithms}

The goal of the first experiment is to characterize the relative performance
of the proposed algorithms. This was done by simulating the algorithms on many
different ensembles and comparing the scores computed using the
performance metric based on exponential scoring defined in
Section~\ref{sec:perf_metric}.

Fig.~\ref{fig:distributions} shows the percentage of simulations for which
each algorithm achieved the highest score for a given ensemble type. This
experiment was conducted using all five applications, with all five types of
ensembles. For each application and ensemble type, 10 random ensembles of 50
workflows each were created. Each ensemble was simulated with all three
algorithms using 10 budgets and 10 deadlines (1,000 simulations per
application, ensemble type, and algorithm, or 75,000 in total). The best scores percentage is
computed by counting the number of times that a given algorithm achieved the
highest score and dividing by 1000. Note that it is possible for multiple
algorithms to get the same high score (to tie), so the numbers do not
necessarily add up to 100\%. The sum is much higher than 100\% in cases where
the dynamic algorithms perform relatively well because DPDS and WA-DPDS, which
are very similar algorithms, often get the same high score.

\begin{figure*}[ht]
    \centering
    \includegraphics[width=\textwidth]{run-finish-variations-test-0-output-distributions}
    \caption{Percentage of high scores achieved by each algorithm on different ensemble types for all five applications. C = Constant ensembles, PS = Pareto Sorted ensembles, PU = Pareto Unsorted ensembles, US = Uniform Sorted ensembles, and UU = Uniform Unsorted ensembles.} 
    \label{fig:distributions}
\end{figure*}

There are several interesting things to notice about
Fig.~\ref{fig:distributions}. The first is that, in most cases, SPSS
significantly outperforms both dynamic algorithms (DPDS and WA-DPDS). This is
attributed to the fact that SPSS is able to make more intelligent scheduling
and provisioning decisions because it has the opportunity to compare different
options and choose the one that results in the best outcome. In comparison,
the dynamic algorithms are online algorithms and are not able to project into
the future to weigh the outcomes of their choices.

The second thing to notice is that, for constant ensembles, the dynamic
algorithms perform significantly better relative to SPSS compared to  other
ensemble types. This is a result of the fact that, since all of the workflows
are of approximately the same size and shape, the choice of which workflow to
execute next has a smaller impact on the final result.

We can also see that the workflow-aware
algorithms (WA-DPDS and SPSS) both perform better in most cases than the
simple online algorithm that uses resource utilization alone to make
provisioning decisions (DPDS). This suggests that there is a significant value
in having information about the structure and estimated runtime of a workflow
when making scheduling and provisioning decisions.

Finally, it is interesting that, for Montage and CyberShake, the
relative performance of SPSS is significantly less than it is for other
applications. We attribute this to the structure of Montage and CyberShake.
The workflows for both applications are very wide relative to their height,
and both have very short-running tasks, resulting  in relatively short
critical paths that makes them look more like bag-of-tasks applications, which are easier
to execute than more structured applications. DPDS and WA-DPDS are able to
pack more of the tasks into the available budget and deadline because there
are a) more choices for where to place the tasks, and b) the different choices
have a smaller impact on the algorithms' ability to execute the workflow
within the constraints. In addition, the short critical paths put SPSS at a
disadvantage. Because of the way SPSS assigns deadlines to individual tasks,
it is prevented from starting workflows late, which prevents it from packing
tasks into the idle VM slots at the end of the schedule.

\subsection{Task Granularity}

In the last experiment we noted that, for Montage and CyberShake, the short
runtimes of their tasks made the dynamic algorithms perform better relative to
SPSS. In order to test this theory we adjusted the granularity of the tasks in
several Montage and CyberShake ensembles to see how this would affect the
relative performance of the algorithms. The granularity adjustment was
achieved by multiplying the runtime of each task by a fixed scaling factor.

Fig.~\ref{fig:stretching} shows the relative performance of the algorithms
as the scaling factor is increased from 1 to 16. The figure only shows results
for the CyberShake application with Pareto unsorted ensembles, but similar
results were found for other ensemble types for both Montage and CyberShake.
Each data point in the figure represents 500 simulations (5 ensembles with 10
budgets and 10 deadlines). The best scores percent was calculated the same way
as it was for Fig.~\ref{fig:distributions}.

The figure shows that, as the scaling factor increases beyond 2, the relative
performance of SPSS surpasses that of the dynamic algorithms. Similar results
were found for other ensemble types for both applications. This result
suggests that, in general, for fine-grained workflows the dynamic algorithms
will produce better scores, and for coarse-grained workflows SPSS will produce
better scores.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.3\textwidth]{stretching_cybershake}
    \caption{Percentage of high scores achieved by each algorithm for Pareto 
    unsorted CyberShake ensembles when task runtime is scaled by a constant 
    factor. A scaling factor of $x$ means that the runtime of tasks in each 
    workflow was multiplied by $x$.}
    \label{fig:stretching}
\end{figure}

\subsection{Inaccurate Task Runtime Estimates}
\label{sec:variances}

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Simulated Cost to Budget]{
        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-cost-all}
        \label{fig:variances-budget}
    }
    \hspace{2cm}
    \subfloat[Ratio of Simulated Makespan to Deadline]{
        \includegraphics[width=0.38\textwidth]{run-finish-variations-test-0-output-makespan-all}
        \label{fig:variances-deadline}
    }
    \caption{Boxplots for budget and deadline ratios when runtime estimate 
    error varies from $\pm$0\% to $\pm$50\% for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:variances}
\end{figure*}

Both of the workflow-aware algorithms rely on estimates of task runtimes to make
better scheduling and provisioning decisions. Our experience suggests that such
assumption is often reasonable, since we can obtain workflow performance
characteristics from preliminary runs~\cite{Bharathi2008,Deelman2005,Juve2010}.
Some applications, like Periodograms~\cite{Vockler2011} even include a
performance model that estimates task runtimes and automatically annotates
workflow description with these data. In practice, however, these
estimates are often inaccurate. Given inaccuate estimates, the question is: How
do errors in task runtime estimates impact the performance of our scheduling and
provisioning algorithms?  To examine this we introduced uniform errors in the
task runtime and observed the behavior of the algorithms in terms of meeting the
desired budget and deadline constraints.

In this experiment the actual runtime of each task is adjusted in the
simulation by adding a random error to the estimated runtime of $\pm p\%$.
Since the sampling is done uniformly, we expect to get just as many
overestimates as underestimates in any given simulation.

Fig.~\ref{fig:variances} shows the results for estimate errors
ranging from $0\%$ to $50\%$. The figure summarizes the outcome of an
extensive suite of 525,000 simulations (10 ensembles of 50 workflows x 5
applications x 5 distributions x 10 budgets x 10 deadlines x 7 error values x
3 algorithms). Box plots show the ratio of the ensemble cost to
budget, and of ensemble makespan to deadline. Whiskers on the plots
indicate maximum and minimum values. The ratio indicates whether the value 
(for example, the simulated cost) exeeded the constraint (the budget). Values
greater than 1 indicate that the constraint was exceeded.

Fig.~\ref{fig:variances-budget} shows the ratio of simulated cost to budget. This
plot illustrates two important algorithm characteristics. First, the dynamic
algorithms very rarely exceeded the budget, even with very large errors. This
indicates that the dynamic algorithms are able to adapt to uncertainties at
runtime to ensure that the constraints are not exceeded, even when the quality
of information available to them is low. Second, unlike the dynamic
algorithms, the static algorithm frequently exceeded the budget constraint; by
large amounts in some cases. This is a result of the fact that the static
algorithm makes all of its decisions ahead of the execution and is not able to adapt
to changing circumstances.

The SPSS plan describes what tasks to execute on which VMs, but not when to
execute them. At runtime, the worklow engine is forced to spend extra money
to extend the runtime of some VMs so that all of the tasks assigned to that
VM can be completed. At the same time, other VMs can be terminated early
because the tasks they were assigned finished earlier than expected. However,
because of dependencies in the workflow, the latter case is less likely to
happen, which causes gaps in the schedule. So the net result is that the
overall cost is increased.

Fig.~\ref{fig:variances-deadline} shows the ratio of simulated makespan to budget.
Interestingly, the deadline constraint is rarely exceeded, even in cases with
very poor quality estimates. For the dynamic algorithms this is a result of the
fact that they can adapt to poor estimates and stop submitting new tasks when
the constraints are reached. Makespan is then computed as the finish time of the
last fully completed workflow from the ensemble. For the static algorithm this
is a result of the way that SPSS schedules workflows, and not of a particularly
clever optimization. The SPSS algorithm tends to schedule workflows early, using
up the budget long before the deadline is reached. This is a consequence of the
deadline distribution function in SPSS, which prevents workflows from starting
late. This is illustrated by the gantt chart in
Fig.~\ref{fig:algorithm-example-spss}, which shows how SPSS tends to pile up
workflows at the beginning of the timeline.
In comparison, the dynamic algorithms tend to spread out workflow starts over
the entire duration as shown in Fig.~\ref{fig:algorithm-example-dpds}. The
consequence of this behavior is that SPSS plans tend to use up the budget but
leave plenty of time before the deadline.
As a result, when the runtime of the plan is increased by introducing errors,
the SPSS plan has some room to expand without exceeding the deadline.

Note that the algorithms have \textit{not} been changed to account for
inaccurate runtime estimates in this experiment. It is likely that better
performance could be achieved if the algorithms were given a hint as to the
accuracy of the task runtimes. Investigating that optimization is left for
future work. 
The goal here is to determine how well the current algorithms are
able to stay within the constraints given inaccurate task runtime estimates.

It is possible that the static algorithm could be modified to add more
breathing room in the schedule to account for situations where the runtime
estimates may be inaccurate. Such a modification would result in slightly
worse scores when the estimates are good, but would improve scores when the
estimates are bad. Alternatively, the SPSS algorithm could be modified to
generate a plan that specifies when to provision and deprovision each VM. In
that case the workflow engine could be prevented from exceeding the budget
and deadline constraints, but would prevent some workflows from finishing,
which may significantly decrease the score. These two modifications are
subjects for future work.

\subsection{Provisioning Delays}
\label{sec:delays}

One important issue to consider when provisioning resources in the cloud is
the amount of time between when a resource is requested, and when it actually
becomes available to the application. Typically these provisioning delays are
on the order of a few minutes, and are highly dependent upon the cloud architecture
and/or the size of the VM image~\cite{Nurmi2008b}.

We assume that resources are billed from the minute that they are requested
until they are terminated.  As a result, provisioning delays have an impact on
both the cost and makespan of an ensemble.

Fig.~\ref{fig:delays} shows the ratios of simulated values to constraints when
the provisioning delay is increased from 0 seconds up to 15 minutes. The
figure summarizes a suite of 105,000 simulations (10 ensembles of 50 workflows
x 5 distributions x 10 budgets x 10 deadlines x 7 error values x 3
algorithms). To reduce the simulation time, only one application, Montage, was
used in this experiment.

As in Fig.~\ref{fig:variances}, the y-axis in each
plot represents the ratio of the simulated value to the constraint value for
the budget constraint or the deadline constraint.

\begin{figure*}[tb]
    \centering
    \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-cost-montage}
        \label{fig:delays-budget}
    }
    \hspace{2cm}
    \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{    
        \includegraphics[width=0.38\textwidth]{run-finish-delays-test-0-output-dagfinish-montage}
        \label{fig:delays-deadline}
    }
    \caption{Boxplots for budget and deadline ratios when provisioning delay 
    varies from 0 seconds to 15 minutes for all three algorithms. Values 
    greater than 1 indicate that the budget/deadline constraint was exceeded.}
    \label{fig:delays}
\end{figure*}

The effect of provisioning delays on workflow performance is similar to that of
inaccurate runtime estimates: when the delays are small, all algorithms are able
to produce results within the constraints, but for larger delays, the dynamic
algorithms are able to adapt to avoid exceeding the constraints while the static
algorithm is not. In the case of delays of more than one minute, which is
typical of what has been observed on academic clouds~\cite{Juve2011} such as
Magellan~\cite{Magellan} and FutureGrid~\cite{FutureGrid}, approximately half of
the SPSS simulations exceeded the budget; in some cases by up to a factor of 2.
In comparison, none of the simulations that used a dynamic algorithm exceeded
the budget or the deadline constraint.

Fig.~\ref{fig:delays_scores} shows the relative performance as a
percentage of the number of high scores achieved by each algorithm as the
provisioning delay increases. This figure shows that, as the delay
increases, the relative performance of the static algorithm increases as
well. This behavior is a result of the fact that the static algorithm is, in
essence, cheating by using more time and money than the dynamic algorithms.
In the previous section we showed how the relative performance of the
algorithms remains the same when a uniform error is applied to the runtime
constraints. In that case, the dynamic algorithms adapted to the error by
rearranging the schedule to accomplish the same amount of work given the
same deadline and budget. In this case, the dynamic algorithms adapt by
performing less work (executing fewer workflows) to remain within the
constraints while accounting for the delays. In both cases, the static
algorithm completed the same amount of work, but did so by exceeding the
constraints.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.5\textwidth]{delays_scores}
  \caption{Percentage of high scores achieved by each algorithm on Montage 
  with uniform unsorted ensembles when provisioning delay varies from 0 
  seconds to 15 minutes.}
  \label{fig:delays_scores}
\end{figure}

This experiment suggests that SPSS is too sensitive to provisioning delays
in its current form to be of practical use in real systems. It is possible
that modifying the SPSS algorithm to account for provisioning delays would
improve its performance on this experiment. In fact, since provisioning
operations are infrequent (because all the algorithms tend to provision
resources for a long time), it is likely that the performance of SPSS could be
improved significantly by simply adding an estimate of the provisioning delay
to its scheduling function. Such an estimate may not have to be particularly
accurate to get good results, and developing an estimate from historical data
should be relatively simple. Testing this idea is left for future work.

\subsection{Task Failures}
\label{sec:failures}

Running workflows consisting of large numbers of tasks on distributed systems
often results in failures. The goal of the next experiment was to assess the
behavior of the algorithms in the presence of task execution failures by
introducing a failure model into the simulator. The model is characterized by
a failure rate $f$ that defines the probability that a task will fail. The
failure time of the task is determined by randomly sampling a value between
the task start time and finish time. If the task fails, it is reported to the
workflow engine, which retries the task until it succeeds. The dynamic
algorithms re-add the task to the priority queue (queue $P$ in
Algorithm~\ref{alg:ds}) so that it can be resubmitted by the scheduler. The
SPSS algorithm immediately  resubmits the failed task to the same VM that was
selected in the plan (to minimize the disruption to the overall plan).

\begin{figure*}[tb]
    \subfloat[Ratio of Simulated Ensemble Cost to Budget]{
        \includegraphics[width=0.38\textwidth]{run-finish-failures-test-0-output-cost-all}
        }
    \hspace{2cm}
    \subfloat[Ratio of Simulated Ensemble Makespan to Deadline]{
        \includegraphics[width=0.38\textwidth]{run-finish-failures-test-0-output-makespan-all}
        }
    \caption{Boxplots for budget and deadline ratios when failure rate varies
    from 0\% to 50\% for all three algorithms. Values greater than 1 indicate 
    that the budget/deadline constraint was exceeded.}
    \label{fig:failures}
\end{figure*}

Fig.~\ref{fig:failures} shows the ratios of simulated values to constraints when
the failures are introduced. The figure summarizes a suite of 525,000
simulations (10 ensembles of 50 workflows x 5 applications x 5 distributions x
10 budgets x 10 deadlines x 7 failure rates x 3 algorithms). As in the previous 
experiments, these results show that high failure rates can degrade the
performance of the static algorithm considerably, while the dynamic algorithms
are able to adapt. Comparing Fig.~\ref{fig:failures} to
Figs.~\ref{fig:variances} and~\ref{fig:delays} one may conclude that failures
are worse than provisioning delays and runtime estimate errors, since their
impact is larger. However, we consider higher failure rates as rare events that
suggest a significant system malfunction or invalid selection of resources.

\subsection{SPSS Planning Time}

Because SPSS involves more complicated logic than the dynamic algorithms and
makes its decisions before execution, it is important to understand what
impact planning time has on the overall execution time.

Fig.~\ref{fig:spss_planning_time} shows the SPSS planning time for ensembles
of 100 workflows with five different workflow sizes: 50, 200, 400, 600, and
800 tasks. The ensembles were generated using a constant distribution equal to
the workflow size desired. Two different applications were used: SIPHT and
CyberShake. Each box summarizes the results of 2000 simulations (2
applications x 10 ensembles x 10 budgets x 10 deadlines).

Fig.~\ref{fig:spss_planning_time} shows that, for small workflows, the SPSS
planning time is reasonable, taking on the order of tens of seconds to a few
minutes. For larger ensembles of large workflows, however, the SPSS planning
time can easily reach 10 minutes. Considering that the largest workflows used
in this experiment are still relatively small (maximum of 800 tasks), and that
real workflows are often much larger (workflows with tens of thousands of
tasks are common, and even workflows with millions of
tasks are possible), it is unlikely that SPSS will be practical for
ensembles of very large workflows.

SPSS considers scheduling each task on the cheapest available slot, which
involves scanning all of the available slots on all of the VMs. Since the number
of available slots, in the worst case, is proportional to the number of tasks
scheduled (because scheduling a task splits an existing slot into at most two
slots: one before the task, and one after), the complexity of SPSS is
$O(n^2)$, where $n$ is the number of tasks in the ensemble. In comparison, the
dynamic algorithms all have a more scalable complexity of $O(n)$. DPDS only
examines the tasks in the workflow once when they are scheduled, and WA-DPDS
does it twice: once in the admission algorithm, and once when they are
scheduled. This makes the dynamic algorithms a better fit for larger workflows
and ensembles even though in some cases they may not produce as good results
as SPSS.

It may be possible to optimize SPSS to reduce its runtime by, for example,
clustering the workflow to increase task granularity, which would decrease the
ratio of planning time to ensemble makespan. It may also be possible to reduce
the complexity of SPSS by employing more sophisticated data structures to
store the available slots. Investigating these topics is left for future work.

\begin{figure}[tb]
    \centering
    \vspace{-6mm}
    \includegraphics[width=0.27\textwidth]{spss_planning_time}
    \vspace{-3mm}
    \caption{Planning time of SPSS algorithm for ensembles of 100 workflows and different workflow sizes.}
    \label{fig:spss_planning_time}
\end{figure}

\section{Conclusions and Future Work}
\label{sec:conclusions}

In this paper we addressed the interesting and important new problem of
scheduling and resource provisioning for scientific workflow ensembles on IaaS
clouds. The goal of this work is to maximize the number of user-prioritized
workflows that can be completed given budget and deadline constraints.

This problem differs from previous work on grid and utility grid scheduling
in that cloud infrastructures provide more control over provisioning, so that
the number of resources can be adjusted according to the requirements of the
application. Therefore the problem space becomes larger; it requires not only
an efficient mapping of tasks to available resources, but also the selection
of the best resource provisioning plan.

Formulating the problem as a maximization of the number of prioritized
workflows completed from the ensemble is also novel and requires workflows to
be admitted or rejected based on their estimated resource demands. We believe
that this bi-constrained problem is highly relevant because such constraints
are commonly imposed on many real-world projects. The approach is also
directly applicable to grid environments that provide resource reservations
and charge service units for resource use.

We developed three algorithms to solve this problem: two dynamic algorithms,
DPDS and WA-DPDS, and one static algorithm, SPSS. The algorithms were
evaluated via simulation on ensembles of synthetic workflows, which were
generated based on statistics from real scientific applications.

The results of our simulation studies indicate that the two algorithms that
take into account the structure of the workflow and task runtime estimates
(WA-DPDS and SPSS) yield better results than the simple priority-based
scheduling strategy (DPDS), which makes provisioning decisions based purely on
resource utilization. This underscores the importance of viewing workflow
ensembles as a whole rather than as individual tasks or individual workflows.

In cases where there are no provisioning delays, task runtime estimates are
good, and failures are rare, we found that SPSS
performs significantly better than both dynamic algorithms. However, when
conditions are less than perfect, the static plans produced by SPSS are
disrupted and it frequently exceeds the budget and deadline constraints. In
comparison, the dynamic algorithms are able to adapt to a wide variety of
conditions, and rarely exceed the constraints even with long delays, poor
estimates, and high failure rates.

In addition, we found that SPSS tends to perform better on coarse-grained
workflows than the dynamic algorithms. For wide, fine-grained workflows, such
as CyberShake and Montage, however, the dynamic algorithms frequently produce
performance that is as good or better than SPSS, because they are better able
to pack the smaller tasks onto idle VMs close to the deadline than SPSS,
which distributes deadlines to tasks in a way that prevents them from
starting late.

For very large workflows and ensembles, we found that the planning time of
the SPSS algorithm is prohibitive. SPSS often took 10 minutes or more to plan
ensembles of 100 workflows with 800 tasks each. This suggests that ensembles
of workflows with tens of thousands of tasks, which are commonly encountered
in real workflow applications, would take many hours to plan using SPSS. This
makes the dynamic algorithms much more attractive for large scale problems.

This study suggests several areas for future work. Our current approach models
data access as part of the task execution time and does not consider data
storage and transfer costs. In the future we plan to extend the application
and infrastructure model to  include the various data storage options
available on clouds. A previous experimental study \cite{Juve2010} suggests
that the data demands of scientific workflows have a large impact on not only
the execution time, but also on the cost of workflows in commercial clouds. 
In dynamic algorithms, it will be interesting to extend our
utilization-based autoscaling with more advanced workflow-aware strategies
based on feedback control.
We also plan to investigate heterogeneous environments that include multiple VM
types and cloud providers, including private and community clouds, which will
make the problem even more complex and challenging.

The results of this study can be applied to develop tools that assist
researchers in planning their large-scale computational experiments. The
estimates of cost, runtime, and number of workflows completed that can be
obtained from both the static algorithms and from the simulation runs,
constitute valuable hints for planning ensembles and evaluating the
associated trade-offs.

\section*{Acknowledgments}
This work was supported by the NSF under award OCI-0943725 (STCI) and
by the EU FP7 projects UrbanFlood (248767) and VPH-Share (269978).

% ACM style
%\bibliographystyle{abbrv}

% IEEE style
\bibliographystyle{IEEEtran}

% This balances the columns by inserting a column break at the given reference
%\IEEEtriggeratref{15}

\bibliography{paper}

%\appendix
%\section{Headings in Appendices}

%\balancecolumns

\end{document}